---
title: PLAN 372 HW 6
author: Nickola Lazarevic
---

```{r}
library(tidyverse)

data = read_csv("TS3_Raw_tree_data.csv")
head(data)


# Check that every City value matches our pattern
all(str_detect(data$City, "^.+, [:upper:]{2}$"))

# See what str_match returns: total match + one column per group
str_match(data$City, "^(.+), ([:upper:]{2})$")

# Assign the groups to new columns
data[,c("city_name", "state")] = str_match(data$City, "^(.+), ([:upper:]{2})$")[,2:3]
head(data)

# Display thw records per state
state_counts <- data %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  arrange(-n)

state_counts

# Data Table
Table_plot <- ggplot(state_counts, aes(x = state, y = n)) +
  geom_col()

view(Table_plot)

#Based on the data there are 13 States total, CA = 4062, WA = 994, ID = 923, HI = 918, FL = 895, IN = 877, SC = 872, CO = 867, NM = 833, NY = 831, NC = 828, AZ = 827 and MN = 760
```


```{r}
# Filter the dataset down to NC and SC 
sc_nc_data <- filter(data, state == "NC" | state == "SC")

# Cities in each state
Table_sc_nc <- sc_nc_data %>% 
  group_by(state, city_name) %>% 
  summarize(n = n()) %>% 
  arrange(state, city_name)

view(Table_sc_nc)

# Based on the table the data was collected from Charlotte in NC and Charleston in SC 
```

```{r}
# Confirm the pattern matches
all(str_detect(sc_nc_data$ScientificName, "^[:alpha:]+ [:alpha:]+"))

# Extract the genus (everything before the first space)
sc_nc_data$genus <- str_match(sc_nc_data$ScientificName, "^([:alpha:]+)")[,2]
head(data)

# Average crown diameter by genus
Avg_crown <- sc_nc_data %>% 
  group_by(genus) %>% 
  summarize(avg_crown = mean(`AvgCdia (m)`, na.rm = TRUE), n = n()) %>% 
  arrange(-avg_crown)

view(Avg_crown)

# Genus with the largest average crown
head(Avg_crown, 1)

# The genus of trees with largest average crown diameter is Quercus and the average crown diameter is 13.62316 meters

```

```{r}
# Average age and crown size by genus
genus_age <- sc_nc_data %>% 
  group_by(genus) %>% 
  summarize(avg_age = mean(Age, na.rm = TRUE),
            avg_crown = mean(`AvgCdia (m)`, na.rm = TRUE),
            n = n()) %>% 
  arrange(-avg_age)

genus_age

# Plot age vs crown for genus grouops
ggplot(genus_age, aes(x = avg_age, y = avg_crown)) +
  geom_point()

#Based on the graph comparing average age of the tree and its average crown size we see that there is a nonlinear relationship meaning that age of the tree does not significantly affect the crown size
```

```{r}
# I used AI to help me with debugging. The use of n() inside summarize() to count the number of observations per group, as in summarize(avg_crown = mean(AvgCdia (m), na.rm = TRUE), n = n()). AI clarified that n() is a tidyverse function that returns the row count for the current group. The use of backticks around column names containing spaces or special characters, as in `AvgCdia (m)`. AI explained that R requires backticks to reference column names that aren't valid identifiers. Overall AI helped me understand how to properly use the n() function and when to use backticks.

```




Place your analysis code in this file; you must not rename it, though you may delete these instructions. Please update the block above with the name of the assignment and with your name.When you push your code to github, it will automatically be rendered for grading. You can see whether the automatic rendering worked in the "Actions" tab of the github repository.

If you use any libraries other than `tidyverse`, `lubridate`, `sf`, `ggplot2`, `here`, `terra`, `ggspatial`, `rJavaEnv`, and `r5r`, add them to the list in the DESCRIPTION file.